---
title: "ARRYTHMIA STUDY"
author: "Ivan Stankov, Kiril Mitchev, Aleksandar Gabrovski, Deian Atanassov"
date: "Tuesday, July 22, 2014"
output: html_document
---

First we need to load our dataset and format in proper way, so we can use it for further modeling.

This loads our matrix of data in a variable called "data" and tells that if there is "?" it is NA value and each value is separated by "," and the decimal symbol is "." 
```{r}
#to load the data properly you must setwd("") where between "" is the path to the folder containing arrhythmia.data
data=read.table("arrhythmia.data",header=F,na.string="?",sep=",", dec=".")
#here we can asign the titles of each variable.
summary(data)
```
Next we will need to manage the NA values. We see that in Var14 there are a lot of NA values. We will change the NA values of this variable with 
the mean of the known values, so we can use it in our modeling. Other variant will be to completely remove it from our model.
```{r}
#the mean of the known values of V14
v14_mean = mean(data$V14,na.rm=T)
#now we replace the NA values with the mean stored in v14_mean
data[is.na(data$V14),14]=v14_mean
```
Now we must look for other variables which have NA values and do the same. After this we are ready to proceed further.
```{r}
data[is.na(data$V11),11] = mean(data$V11,na.rm=T)
data[is.na(data$V12),12] = mean(data$V12,na.rm=T)
data[is.na(data$V13),13] = mean(data$V13,na.rm=T)
data[is.na(data$V15),15] = mean(data$V15,na.rm=T)
```
Removing the variables with no variance and NAs
```{r}
library(caret)
zeroVar=c(nearZeroVar(data))
data=data[,-zeroVar]# removing the Variables with zero
data$V14=NULL #removing the Variable with NAs
```
Now we plot some boxplots over data to briefly visualise it. 
```{r}
# there must be more rational way to do this!!!
boxplot(data[1:20])
boxplot(data[21:40])
boxplot(data[41:60])
boxplot(data[61:80])
boxplot(data[81:100])
boxplot(data[101:120])
boxplot(data[121:140])
boxplot(data[141:148])

```
V2 is Sex - we make it as.factor
```{r}
data$V2=as.factor(data$V2)
```
Box-Cox transformation (for the paper include formula on page 32)
```{r}
data1=preProcess(data[,-c(2,148)], method = "BoxCox")
data11=predict(data1,data)
```
Creating dihotomous response variable
1 for normal; 0 for diseased
```{r}
response=rep(0,nrow(data11))
for (i in 1:length(response)){
  if (data11$V280[i]==1){response[i]=1}
  else {response[i]=0}
}

#data11 becomes data, and is without the last variable
data11$V280=response
data=data11
remove(data11)



# PRINCIPAL COMPONENT ANALYSIS

# Creating data without the dummy variables
data2=data
data2$V2=NULL

# Looking for differences in variable means and variances
apply(data2,2,mean)
apply(data2,2,var)

# Running the initial uncalibrated principal components algorhythm
PC=prcomp(data2 , scale =TRUE)
summary(PC)
names(PC)
PC$center
PC$rotation
dim(PC$x)

# Determine the part of variance explained by the principal components
PC$sdev
PCvar=PC$sdev^2
PCvar
PCvarexp=PCvar/sum(PCvar)
PCvarexp

# Plot the principal components to find the optimal number to retain
plot(PCvarexp ,xlab="Principal Component",ylab="Proportion of Variance Explained",ylim=c(0,1) ,type="b")
plot(cumsum(PCvarexp),xlab="Principal Component",ylab ="Cumulative Proportion of Variance Explained",ylim=c(0,1),type="b")

# Running a calibrated principal component algorhythm to limit the components to the significant ones
PC2=prcomp(data2 , scale =TRUE,tol=0.33)
summary(PC2)

# K_MEANS CLUSTERING

# Creating new dataset without response variable
data3=data
data3$V280=NULL

# Determine number of clusters that minimizes the within-group distances

wss <- (nrow(data3)-1)*sum(apply(data3,2,var))
for (i in 2:20) wss[i] <- sum(kmeans(data3,
   centers=i)$withinss)
plot(1:20, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares") 

# K-Means Cluster Analysis

# 16 cluster solution
KM=kmeans(data3,16,nstart=20)

# get cluster means
aggregate(data3,by=list(KM$cluster),FUN=mean)

# append cluster assignment
data4=data.frame(data3,KM$cluster)


