---
title: "ARRYTHMIA STUDY"
author: "Ivan Stankov, Kiril Mitchev, Aleksandar Gabrovski, Deian Atanassov"
date: "Tuesday, July 22, 2014"
output: html_document
---

First we need to load our dataset and format in proper way, so we can use it for further modeling.

This loads our matrix of data in a variable called "data" and tells that if there is "?" it is NA value and each value is separated by "," and the decimal symbol is "." 
```{r}
#to load the data properly you must setwd("") where between "" is the path to the folder containing arrhythmia.data
data=read.table("arrhythmia.data",header=F,na.string="?",sep=",", dec=".")
#here we can asign the titles of each variable.
summary(data)
```
Next we will need to manage the NA values. We see that in Var14 there are a lot of NA values. We will change the NA values of this variable with 
the mean of the known values, so we can use it in our modeling. Other variant will be to completely remove it from our model.
```{r}
#the mean of the known values of V14
v14_mean = mean(data$V14,na.rm=T)
#now we replace the NA values with the mean stored in v14_mean
data[is.na(data$V14),14]=v14_mean
```
Now we must look for other variables which have NA values and do the same. After this we are ready to proceed further.
```{r}
data[is.na(data$V11),11] = mean(data$V11,na.rm=T)
data[is.na(data$V12),12] = mean(data$V12,na.rm=T)
data[is.na(data$V11),13] = mean(data$V13,na.rm=T)
data[is.na(data$V15),11] = mean(data$V15,na.rm=T)
```
Removing the variables with no variance and NAs
```{r}
library(caret)
zeroVar=c(nearZeroVar(data))
data=data[,-zeroVar]# removing the Variables with zero
data$V14=NULL #removing the Variable with NAs
```
Now we plot some boxplots over data to briefly visualise it. 
```{r}
# there must be more rational way to do this!!!
boxplot(data[1:20])
boxplot(data[21:40])
boxplot(data[41:60])
boxplot(data[61:80])
boxplot(data[81:100])
boxplot(data[101:120])
boxplot(data[121:140])
boxplot(data[141:148])

```
V2 is Sex - we make it as.factor
```{r}
data$V2=as.factor(data$V2)
```
Box-Cox transformation (for the paper include formula on page 32)
```{r}
data1=preProcess(data[,-c(2,148)], method = "BoxCox")
data11=predict(data1,data)
```
Creating dihotomous response variable
1 for normal; 0 for diseased
```{r}
response=rep(0,nrow(data11))
for (i in 1:length(response)){
  if (data11$V280[i]==1){response[i]=1}
  else {response[i]=0}
}

#data11 becomes data, and is without the last variable
data11$V280=NULL
data=data11
remove(data11)












